\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-04-19 forwards-and-backwards.tex
edits until defn 6.1

Putting these together,
the level $u$ of an individual at $x$
evolves according to
$$
\dot{u}
=
    - \frac{u}{\lambda} \left(\lambda - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) 
    +
    \mu(x,\eta) u .
$$
We shall write 
$$
    b(x, \eta)
    :=
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    -
    \mu(x,\eta),
$$
which captures the local net difference between reproduction and death,
i.e., the deviation from criticality of the branching mechanism.
(This differs from $F = \theta(r\gamma - \mu)$ defined above
both by the factor of $\theta$ and
because $b$ contains the adjustment to fecundity obtained by
averaging over establishment probabilities in the locations that offspring disperse to.)
We can then rewrite the differential equation 
governing the how each individual's level evolves as
\begin{align}
\dot{u}
    &=
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    \left\{
        -\frac{u}{\lambda}\left(\lambda - u\right)
        + u
    \right\}
    -
    b(x,\eta)u
    \nonumber \\
    &=
    \lambda^{-1} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
    -
    b(x, \eta)u
    . \label{differential equation for level}
\end{align}

Now, we can write down the generator for $(\lp_t)_{t \ge 0}$,
the lookdown process.
In what follows, we will write sums (and, products) over ``$(x, u) \in \xi$''
to mean as a sum over the (location, level) pairs of each individual in the population.
Test functions for $\lp$ then take the form
\begin{equation}
\label{test functions}
f(\lp)=\prod_{(x,u)\in \lp}g(x,u)=\exp\left(\int \log g(x,u)\lp(dx, du)\right),
\end{equation}
where
$g(x,u)$ is differentiable in $u$ and 
smooth in $x$.
We will also assume that $0\leq g(x,u) \leq 1$ for all $u\in [0,\lambda]$,
and $g(x,u)\equiv 1$ for $u\geq \lambda$.
In the expressions that follow,
we shall often see one or more factor of $1/g(x,u)$;
it should be understood that if $g(x,u)=0$,
then it simply cancels 
the corresponding factor in $f(\lp)$.

@@

First consider the terms in the generator that come from birth events.
Recall that when a birth successfully establishes,
a new level is generated above the parent's level,
and this new level is assigned to either the offspring or the parent.
Since the probability of each is 1/2,
the contribution of birth to the generator is
\begin{align*}
f(\lp)
&\mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \frac{\theta}{\lambda} \gamma(x, \eta)
    \int_u^\lambda
    \int_{\IR^d}
    \left(
    \frac{1}{2}
    \bigg\{
            g(y, u_1)
        + \frac{ g(y, u) g(x, u_1) }{ g(x, u) }
    \bigg\}
        - 1
    \right)
    r(y, \eta) q(x, dy)
    \\ \begin{split}
&=
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \gamma(x, \eta)
    \bigg\{
        \frac{1}{2 \lambda}
        \int_u^\lambda
        g(x, u_1) du_1
        \frac{
            \theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q(x, dy)
        }{
            g(x, u)
        }
    \\ & \qquad \qquad \qquad {}
        + \frac{\theta}{\lambda}
        \int_u^\lambda \int_{\IR^d}
        \left( \frac{g(y, u_1) + g(x, v_1)}{2} - 1 \right)
        r(y, \eta) q(x, dy)
    \bigg\}
    .
    \end{split}
\end{align*}
Here, $u_1$ is the new level and $y$ is the offspring's location,
and so the two terms in the integral correspond to the two situations:
in the first, we have added an individual at $(y, u)$,
while in the second, we replace an individual at $(x, u)$
by one at $(x, u_1)$ and another at $(y, u)$.

The remaining term in the generator is due to the motion of particles' levels:
\begin{align*}
    f(\lp)
    \mapsto
    f(\lp)
    \theta
    \sum_{(x, u) \in \lp}
    \left(
        \lambda^{-1} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
        -
        b(x, \eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)} .
\end{align*}

The generator of $\lp$ applied to functions of this form
is therefore
\begin{eqnarray}
&&
A f(\lp ) \nonumber \\
&&\quad
    = f(\lp) \,
    \sum_{(x,u) \in \lp}\,
        2 \lambda^{-1} \gamma(x, \eta)
    \int_u^\lambda \Bigg(
        \frac{1}{2} \frac{
            \int_{\mathbb{R}^d} g(y,u) r(y,\eta) q(x, dy) g(x,v_1)
        }{
            g(x,u)
        }
\label{multgen3}\\
&&\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad {}
        + \frac{1}{2} \int_{\mathbb{R}^d} g(y,v_1) r(y,\eta) q(x,dy)
        - 1
    \Bigg) dv_1
\nonumber\\
&&\quad \quad
    + f(\lp) \left(
        \lambda^{-1} \theta
        \gamma(x,\eta)
        \int_{\IR^d} r(y, \eta q(x, dy)
        u^2
        -
        b(x,\eta)u
    \right)
    \frac {\partial_u g(x,u)}{g(x,u)}
\nonumber\\
&&\quad
    = f(\lp) \,
    \sum_{(x,u)\in\lp}\,
    2 \gamma(x, \eta) \frac{\theta}{\lambda}
    \int_u^\lambda \Bigg(
        \frac{1}{2} \frac{
            \int_{\mathbb{R}^d}( g(y,u) - g(x,u) ) r(y, \eta) q(x,dy) g(x,v_1)
        }{
            g(x,u)
        }
\non\\
&&\qquad\qquad\qquad\qquad\qquad\qquad\qquad {}
        + \int_{\mathbb{R}^d} \left(
            \frac{ g(y,v_1) + g(x,v_1) }{ 2 }
            -
            1
        \right)
        r(y, \eta) q(x,dy)
    \Bigg) dv_1
\nonumber\\
&&\qquad\qquad {}
    +
    f(\lp) \sum_{(x,u)\in\lp}\,
    \left(
        \lambda^{-1}
        \gamma(x,\eta)
        \int_{\IR^d} r(y, \eta) q(x, dy)
        u^2
        -
        b(x,\eta) u
    \right)
    \frac{ \partial_u g(x,u) }{ g(x,u) }
.\nonumber
\end{eqnarray}
The two terms under the integral correspond, respectively,
to either the parent or child being reassigned the new level $v_1$
during a birth event.
In the first case, the parent at location $x$,
which had level $u$ before the event, is reassigned to level $v_1$,
and the child is inserted at level $u$ and location $y$;
in the second the parent retains its level,
and offspring is inserted at level $v_1$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Scaling of the Pre-limit Lookdown Model}

To scale our lookdown model to be consistent with
the spatial model described in \S \ref{sec: Model and main results},
we will let $\lambda$ takes on the role of $N$
in \S \ref{sec: Model and main results},
i.e. each individual has mass $1/\lambda$
in the scaling limit.
Time is once again scaled by $\theta$,
and the birth, death and establishment rates 
depend on local densities given by
mollification over Gaussian kernels with radius
$\epsilon_{\gamma}$, $\epsilon_{F}$ and $\epsilon_{r}$.

Explicitly, in the $\theta$-th stage of re-scaling,
an individual has mass $\frac{1}{\lambda} := \frac{1}{N}$
and a level between $[0,N]$;
the established birth rate $\gamma(x,\eta)$ is
$$\theta\gamma^{\mathfrak{m}}_{\theta}(x, \eta)
:=~\theta\gamma(x, \rho_{\gamma}*\eta(x)) \int r(y, \rho_{r}*\eta(y))q_{\theta}(x,dy);$$
the local branching mechanism $b(x,\eta)$ is 
$$b_{\theta}(x, \eta):=~ \theta \bigg\{ \gamma(x, \rho_{\gamma}*\eta(x)) \int r(y, \rho_{r}*\eta(y))q_{\theta}(x,dy)-\mu(x, \eta)\bigg\};$$
and the dispersal kernel $q^{\mathfrak{m}}(x,\eta, dy)$ is 
$$q^{\mathfrak{m}}_{\theta}(x,\eta,dy) 
:=~ \frac{r(y,\rho_{r}*\eta(y)) q_{\theta}(x,dy)}{\int_{\mathbb{R}^d} r(y,\rho_{r}*\eta(y)) q_{\theta}(x,dy)}.$$

We can now define our pre-limit explicitly as a solution to a Martingale Problem.
\comment{TODO: this is a bit redundant with the above.
Instead, maybe (a) describe the process in words; (b) give this definition;
then (c) explain why it's the generator, so we only write down the generator once?}

% % % % % % % %
\begin{definition}[Martingale Problem Characterisation]
\label{Def: Rescaled Lookdown Models}

For test measures $\xi$ taking value in $\mathcal{M}_F(\mathbb{R}^d \times [0,N])$,
the set of finite measures on $\mathbb{R}^d \times [0,N]$,
we define
$$\eta:= \frac{1}{N}\xi(\cdot  \times [0, N])=
\frac{1}{N}\sum_{(x,u) \in \xi}\delta_x$$
to be its re-scaled spatial projection.

We define $(\xi^{\theta}_t)_{t \geq 0}$
to be the (unique) solution
to the Martingale Problem
with initial condition $\xi_0$
and generator $A^{\theta}$ that satisfies

\begin{eqnarray}\label{Eq: Prelimit Lookdown Generator}
&&A^{\theta}f(\xi ) \\
&&\quad =f(\xi )\,\sum_{(x,u)\in\eta}\,2 \gamma^{\mathfrak{m}}_{\theta}(x, \eta)\Bigg\{ \frac 12 N^{-1}\int_u^{
N}g(x,v_1)dv_1\times \frac { \theta\int_{\mathbb{R}^d}(g(y,u)-g(x,u))q^{\mathfrak{m}}_{\theta}(x,\eta,dy) }{g(x,u)}\non\\
&&\qquad\qquad\qquad\qquad\qquad\qquad\qquad +N^{-1}\theta\int_u^{
N}\int_{\mathbb{R}^d}\left(\frac{g(y,v_1)+g(x,v_1)}{2}-1\right)q^{\mathfrak{m}}_{\theta}(x,\eta,dy)dv_1\nonumber\Bigg\}\\
&&\qquad\qquad
+f(\eta )\sum_{(x,u)\in\xi}\,\left(N^{-1} \theta \gamma^{\mathfrak{m}}_{\theta}(x,\eta) u^2 -b_{\theta}(x,\eta)u\right)\frac {\partial_u g(x,u)}{g(x,u)}.\nonumber
\end{eqnarray}
for all smooth functions $f \in C^{2}_{0}(\mathbb{R}^d)$. 
\end{definition}

\comment{Remark on the Remark: this is more than one remark.
I'm not sure it needs to be in a Remark block: how about just a few paragraphs?}

In Appendix \ref{sec: Markov Mapping Theorem Application},
we will show that
if we define $\Gamma (d\textbf{u})$
as the probability kernel that corresponds to
assigning each particle an i.i.d. uniform $[0,N]$-level,
and for any $\xi \in \mathcal{N}_f(\mathbb{R}^d \times [0,\infty))$, we denote
$$\hat{f}(\xi)=\int f(\xi) \Gamma (d\textbf{u})$$ 
to be the spatial test functions with averaged level,
then the process
\begin{equation}
\begin{aligned}
M^{\theta}_t:=&\hat{f}(\xi_t)-\hat{f}(\xi_0)-\int_{0}^{t}\int   A^{\theta}f(\xi_s)\Gamma(d\textbf{u})ds\\
:=&\hat{f}(\xi_t)-\hat{f}(\xi_0)-\int_{0}^{t}   \Pgen^{\theta}\hat{f}(\xi_s)ds
\end{aligned}    
\end{equation}
is a martingale.

@@

Therefore, we use the same notation $\eta$
in our spatial model
to denote the spatial projection of $\xi$,
and we reserve the notation $\xi$
strictly for the spatial-level process.

>>>>

Putting these together,
the level $u$ of an individual at $x$
evolves according to
$$
\dot{u}
=
    - \frac{u}{\lambda} \left(\lambda - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) 
    +
    \mu(x,\eta) u .
$$
We shall write 
$$
    b(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    -
    \mu(x,\eta)
    \right) ,
$$
which captures the local net difference between reproduction and death,
i.e., the deviation from criticality of the branching mechanism.
(This differs from $F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu(x,\eta))$ defined above
because $b$ contains the adjustment to fecundity obtained by
averaging over establishment probabilities in the locations that offspring disperse to.)
\comment{TODO: make this explanation better.}
We can then rewrite the differential equation 
governing the how each individual's level evolves as
\begin{align}
\dot{u}
    &=
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    \left\{
        -\frac{u}{\lambda}\left(\lambda - u\right)
        + u
    \right\}
    -
    b(x,\eta)u
    \nonumber \\
    &=
    \lambda^{-1} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
    -
    b(x, \eta) u / \theta
    . \label{differential equation for level}
\end{align}

Now, we can write down the generator for $(\lp_t)_{t \ge 0}$,
the lookdown process.
In what follows, we will write sums (and, products) over ``$(x, u) \in \xi$''
to mean as a sum over the (location, level) pairs of each individual in the population.
Test functions for $\lp$ then take the form
\begin{equation} \label{eqn:test_functions}
f(\lp)=\prod_{(x,u)\in \lp}g(x,u)=\exp\left(\int \log g(x,u)\lp(dx, du)\right),
\end{equation}
where
$g(x,u)$ is differentiable in $u$ and 
smooth in $x$.
We will also assume that $0\leq g(x,u) \leq 1$ for all $u\in [0,\lambda]$,
and $g(x,u)\equiv 1$ for $u\geq \lambda$.
In the expressions that follow,
we shall often see one or more factor of $1/g(x,u)$;
it should be understood that if $g(x,u)=0$,
then it simply cancels 
the corresponding factor in $f(\lp)$.

@@

First consider the terms in the generator that come from birth events.
Recall that when a birth successfully establishes,
a new level is generated above the parent's level,
and this new level is assigned to either the offspring or the parent.
Since the probability of each is 1/2,
the contribution of birth to the generator is
\begin{align}
f(\lp) \nonumber
&\mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \frac{\theta}{\lambda} \gamma(x, \eta)
    \int_u^\lambda
    \int_{\IR^d}
    \left(
    \frac{1}{2}
    \bigg\{
            g(y, u_1)
        + \frac{ g(y, u) g(x, u_1) }{ g(x, u) }
    \bigg\}
        - 1
    \right)
    r(y, \eta) q(x, dy)
    \\
    \begin{split} \label{eqn:birth_generator}
&=
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \gamma(x, \eta)
    \bigg\{
        \frac{1}{2 \lambda}
        \int_u^\lambda
        g(x, u_1) du_1
        \frac{
            \theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q(x, dy)
        }{
            g(x, u)
        }
    \\ & \qquad \qquad \qquad {}
        + \frac{\theta}{\lambda}
        \int_u^\lambda \int_{\IR^d}
        \left( \frac{g(y, u_1) + g(x, v_1)}{2} - 1 \right)
        r(y, \eta) q(x, dy)
    \bigg\}
    .
    \end{split}
\end{align}
Here, $u_1$ is the new level and $y$ is the offspring's location,
and so the two terms in the integral correspond to the two situations:
in the first, we have added an individual at $(y, u)$,
while in the second, we replace an individual at $(x, u)$
by one at $(x, u_1)$ and another at $(y, u)$.

The remaining term in the generator is due to the motion of particles' levels:
\begin{align} \label{eqn:level_generator}
    f(\lp)
    \mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    \left(
    \theta
        \lambda^{-1} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
        -
        b(x, \eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)} .
\end{align}

\comment{TODO: consider whether this is where we stick in
    these explicit definitions of $\gamma$ and set $\lambda=N$.}
To scale our lookdown model to be consistent with
the spatial model described in \S \ref{sec: Model and main results},
we will let $\lambda$ takes on the role of $N$
in \S \ref{sec: Model and main results},
i.e. each individual has mass $1/\lambda$
in the scaling limit.
Time is once again scaled by $\theta$,
and the birth, death and establishment rates 
depend on local densities given by
mollification over Gaussian kernels with radius
$\epsilon_{\gamma}$, $\epsilon_{F}$ and $\epsilon_{r}$.

Explicitly, in the $\theta$-th stage of re-scaling,
an individual has mass $\frac{1}{\lambda} := \frac{1}{N}$
and a level between $[0,N]$;
the birth rate is
$$
    \gamma(x, \eta)
    :=
    \gamma(x, \smooth{\gamma}\eta(x)),
$$
the establishment probability is
$$
    r(x, \eta)
    :=
    r(x, \smooth{r}\eta(x)),
$$
the local branching mechanism $b(x,\eta)$ is 
$$
    b_{\theta}(x, \eta)
    :=
    \theta \left(
        \gamma(x, \smooth{\gamma}\eta(x))
        \int r(y, \smooth{r}\eta(y)) q_{\theta}(x,dy)
        -
        \mu(x, \eta)
    \right) .
$$
and the dispersal kernel is $q_\theta$.

We can now define the spatial-level process
explicitly as a solution to a Martingale Problem,
whose generator is just the sum of
\eqref{eqn:birth_generator} and \eqref{eqn:level_generator}.

% % % % % % % %
\begin{definition}[Martingale Problem Characterisation]
    \label{Def: Rescaled Lookdown Models}
We define $(\lp^{\theta}_t)_{t \geq 0}$
to be the (unique) solution
    \comment{(I'm not sure why unique is in parentheses here?)}
to the Martingale Problem
with initial condition $\lp_0 \in \mathcal{M}_F(\mathbb{R}^d \times [0,N])$
and generator $A^{\theta}$ that satisfies
\begin{equation*}
\begin{split}
& A^{\theta}f(\lp ) \\
&\quad =
    f(\lp)
    \,\sum_{(x,u)\in \lp}\,
    2 \gamma(x, \eta)
    \Bigg\{ \frac 12 N^{-1}\int_u^N g(x,u_1) du_1
            \times
            \frac{
                \theta \int_{\mathbb{R}^d}
                (g(y,u) - g(x,u))
                r(y, \eta) q_{\theta}(x,dy)
            }{ g(x,u) }
        \\
    &\qquad\qquad\qquad\qquad\qquad\qquad\qquad {} +
        N^{-1}\theta\int_u^N
        \int_{\mathbb{R}^d}\left(
            \frac{ g(y,u_1) + g(x,u_1) }{ 2 } - 1
        \right)
        r(y, \eta) q_{\theta}(x,dy)
        du_1
    \Bigg\}\\
    &\qquad\qquad {} +
    f(\lp) \sum_{(x,u)\in\lp}\,
    \left(
        \theta N^{-1} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) u^2 -b_{\theta}(x,\eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)}
    ,
\end{split}
\end{equation*}
for all smooth functions $g \in C^{2}_{0}(\mathbb{R}^d)$,
where $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$ as defined in \eqref{eqn:test_functions},
and $\eta(\cdot) = \lp(\cdot \times [0, N]) / N$ is the spatial pushforward of $\lp$,
as before.
\end{definition}

In Appendix \ref{sec: Markov Mapping Theorem Application},
we will show that
if we define $\Gamma (d\textbf{u})$
as the probability kernel that corresponds to
assigning each particle an i.i.d. uniform $[0,N]$-level,
and for any $\lp \in \mathcal{M}_f(\mathbb{R}^d \times [0,\infty))$, we denote
$$\hat{f}(\lp)=\int f(\lp) \Gamma (d\textbf{u})$$ 
to be the spatial test functions with averaged level,
then the process
\begin{equation}
\begin{aligned}
M^{\theta}_t:=&\hat{f}(\lp_t)-\hat{f}(\lp_0)-\int_{0}^{t}\int   A^{\theta}f(\lp_s)\Gamma(d\textbf{u})ds\\
:=&\hat{f}(\lp_t)-\hat{f}(\lp_0)-\int_{0}^{t}   \Pgen^{\theta}\hat{f}(\lp_s)ds
\end{aligned}    
\end{equation}
is a martingale.

@@
